---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# kaggle-loans-for-house-purchase-in-austria

<!-- badges: start -->
<!-- badges: end -->

This repository documents how the data for the Kaggle competition [Loans to households for house purchase in Austria](https://www.kaggle.com/competitions/loans-to-households-for-house-purchase-in-austria) was generated.


```{r, message=FALSE}
library(dplyr)
library(ecb)
library(lubridate)
library(readxl)
library(tidyr)
library(zoo)
```

## ECB data

We use data from ECB's data portal on new loans for house purchase issued by Austrian banks to private households.

```{r, message=FALSE}
vol <- get_data("MIR.M.AT.B.A2C.A.B.A.2250.EUR.P") %>%
  select(obstime, obsvalue) %>%
  rename(date = obstime, crdt = obsvalue) %>%
  na.omit()

intr <- get_data("MIR.M.AT.B.A2C.A.R.A.2250.EUR.P") %>%
  select(obstime, obsvalue) %>%
  rename(date = obstime, r = obsvalue) %>%
  na.omit()

result <- full_join(vol, intr, by = "date") %>%
  arrange(date) %>%
  mutate(date = as.Date(paste0(date, "-01")))
```

## KIM-V: Borrower-base measures

The Austria Financial Market Authority introduced legally binding borrower-based measures to address systemic risks from the national residential real estate sector. The measures were legally binding between August 2022 and June 2025.

```{r}
result <- result %>%
  mutate(kim = as.numeric(date >= "2022-08-01" & date <= "2025-06-01"))
```

## European business and consumer survey

```{r, message=FALSE, warning=FALSE, error=FALSE, results='hide'}

suffix <- "consumer_total_sa_nace2"
tmpfile <- tempfile(tmpdir = tmpdir <- tempdir())

# Try download for current month
curr_date <- Sys.Date()
curr_month <- lubridate::month(curr_date)
curr_month <- ifelse(nchar(curr_month) == 1, paste0("0", curr_month), curr_month)
curr_year <- substring(lubridate::year(curr_date), 3, 4)
curr_month <- paste0(curr_year, curr_month)
try(download.file(paste0("https://ec.europa.eu/economy_finance/db_indicators/surveys/documents/series/nace2_ecfin_", curr_month, "/", suffix, ".zip"),
                  destfile = tmpfile),
    silent = TRUE)
sdmx_files <- unzip(tmpfile, exdir = tmpdir)

# Download failed try it with download of one month earlier
if (length(sdmx_files) == 0) {
  curr_date <- lubridate::floor_date(Sys.Date(), "month") - 1
  curr_month <- lubridate::month(curr_date)
  curr_month <- ifelse(nchar(curr_month) == 1, paste0("0", curr_month), curr_month)
  curr_year <- substring(lubridate::year(curr_date), 3, 4)
  curr_month <- paste0(curr_year, curr_month)
  try(download.file(paste0("https://ec.europa.eu/economy_finance/db_indicators/surveys/documents/series/nace2_ecfin_", curr_month, "/", suffix, ".zip"),
                    destfile = tmpfile))
}

zipped_files <- unzip(tmpfile, exdir = tmpdir)

monthly_data <- read_excel(zipped_files, "CONSUMER MONTHLY", na = "NA")
names(monthly_data)[1] <- "date"

quarterly_data <- read_excel(zipped_files, "CONSUMER QUARTERLY", na = "NA")
names(quarterly_data)[1] <- "date"
quarterly_data <- quarterly_data %>%
  mutate(date = as.Date(as.yearqtr(date, "%Y-Q%q")),
         date = ceiling_date(date, "quarter") - 1)

survey <- bind_rows(monthly_data,
                 quarterly_data) %>%
  pivot_longer(cols = -c("date")) %>%
  filter(!is.na(value)) %>%
  mutate(date = floor_date(as.Date(date), "month"),
         ctry = substring(name, 6, 7),
         name = substring(name, 13, nchar(name)),
         question = regmatches(name, regexpr("[^.]*", name))) %>%
  filter(ctry == "AT") %>%
  select(date, question, value) %>%
  mutate(question = paste0("survey_", question)) %>%
  pivot_wider(names_from = "question", values_from = "value")

result <- left_join(result, survey, by = "date")
```

## Seasonality

```{r, message=FALSE}
result <- result %>%
  mutate(season_jan = as.numeric(substring(date, 6, 7) == "01"),
         season_feb = as.numeric(substring(date, 6, 7) == "02"),
         season_mar = as.numeric(substring(date, 6, 7) == "03"),
         season_apr = as.numeric(substring(date, 6, 7) == "04"),
         season_may = as.numeric(substring(date, 6, 7) == "05"),
         season_jun = as.numeric(substring(date, 6, 7) == "06"),
         season_jul = as.numeric(substring(date, 6, 7) == "07"),
         season_aug = as.numeric(substring(date, 6, 7) == "08"),
         season_sep = as.numeric(substring(date, 6, 7) == "09"),
         season_oct = as.numeric(substring(date, 6, 7) == "10"),
         season_nov = as.numeric(substring(date, 6, 7) == "11"),
         season_dec = as.numeric(substring(date, 6, 7) == "12"))
```

## Save data

```{r}
write.csv(result, file = "loans-for-house-purchase-in-austria.csv", row.names = FALSE)
```

